Comparing variations in sediment transport capacity between subglacial and subaerial systems

### 2023-03-20 theory.tex
Pushing back to mac

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\marie}[1]{{\textbf{\color{green}Marie says:} \color{green} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
  While water velocity in the glacier system varies by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$.
  Similarly, shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$. This increased variability  in shear stress between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.

  Changing channel geometry accommodates evolving water discharge in the subaerial system (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only be accommodated by an increase in velocity (Equation~\ref{eq:DW}).
  The range of values between width-integrated shear stress in the subaerial system increases to $0.44$, whereas, variability in the glacial system lies at $0.83$ (Figure~\ref{fig:range}).
  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs.
  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates a portion of the sediment transport capacity.

  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmma). $Q_w^*$ in Equation ~\ref{eq:DW} is determined as the midpoint in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

@@

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones (Figure~\ref{fig:range} g, h).
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S \ref{fig:gammas}).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).

>>>>

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO #1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\marie}[1]{{\textbf{\color{green}Marie says:} \color{green} #1} }
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

  Across the hydrological forcing, variable water discharge results in a substantial increase in the range of water velocities in the glacial system compared to the subaerial system (Figure~\ref{fig:model_outs}).
  Water velocity in the glacier system vary by  $0.5$ (Equation~\ref{eq:var}) over the discharge values, in the subaerial system the water velocity only varies by $0.11$ (Figure~\ref{fig:model_outs}\,e).
  Shear stress at the glacier bed varies by $0.75$ across the range of water discharge values, whereas shear stress across the subaerial system varies by roughly $0.21$ (Figure~\ref{fig:model_outs}\,f). The increased variability in shear stress relative to water velocity between the two systems occurs as the water velocity is raised to the power of $2$ in Equation~\ref{eq:tau}.

  In the subaerial systems, changing channel geometry accommodates evolving water discharge  (Equation~\ref{eq:wcf}), whereas in the subglacial case, increased water flux can only result in changing water velocity (Equations~\ref{eq:v}~ and ~\ref{eq:DW}).
  As a result, in the subaerial system, variability in the shear stress $\tau_f$ and width-integrated shear stress $\tau_{tf}$ increases from $0.21$ to $0.44$. Variability in width-integrated shear stress  in the glacial system $\tau_{tg}$ is $0.83$ (Figure~\ref{fig:range}).
  The reduced variability in width-integrated shear stress results from the additional width across the subaerial channel that accommodates some variations in water discharge.
  Indeed, the relative  difference in width-integrated shear stress is less between the subaerial and glacier cases, compared to the water velocity and shear stress outputs (Figure~\ref{fig:model_outs}\,e--g).

  Next, the parameterizations are run $25,000$ times with randomly selected channel geometries ($k=(2,6)$, $\beta=(\frac{\pi}{10},\pi)$), friction factors ($f_i=(0.001,20)$, $f_p=(0.001,20)$) and water discharges (maximum and minimum values selected between $5$ to $500$ \,\mmms). $Q_w^*$ in Equation ~\ref{eq:DW} is determined from the mean in the water discharge range. Parameter combinations are only selected if their maximum and minimum velocities for both the subglacial and subaerial systems are between $0.2$ and $3$\,\unit{m}\,\unit{s}$^{-1}$, resulting in roughly $10,600$ adequate parameter combinations.

@@

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}). 
\ian{extra sentence is needed here}
